A survey of HTTP caching options on the open Semantic Web

### 2015-01-16 sanity/paper1.tex
Martin proofread
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@997 43ef1190-90fd-4dd7-a769-d0121f02b80e

Caching has been given a prominent place in the foundational documents
of the World Wide Web. Out of the 6 documents that make up the HTTP
1.1 standard, RFC7234 \cite{rfc7234} is entirely devoted to the topic,
and RFC7232 \cite{rfc7232} defines conditional requests, which is also
important when constructing caches. As the former notes:
\begin{quote} 
  The goal of caching in HTTP/1.1 is to significantly improve
  performance by reusing a prior response message to satisfy a current
  request.
\end{quote}
Furthermore, the Architecture of the World Wide Web
\cite{Jacobs:04:AWW} discusses caching throughout, and the definition
of the Representational State Transfer (REST) architectural style
\cite{Fielding_2000_Architectural-Styles} is partly motivated from the
requirement to implement efficient caching. We also note that caching
in the Internet infrastructure, through so-called Content Delivery
Networks, is both a large business area and something that could
provide great value to the Semantic Web community.

@@

We note that caching is not only useful for long-living information
resources, even though that may be the most important. If a resource
is frequently requested, it may make sense to cache it even though it
may be fresh for only a very short period.

Caching may be deployed at several different levels: A HTTP cache may
be in a reverse proxy close to the server, in which case it may have
much in common with a conventional database cache. It may also be
anywhere between a server and a client, in which case it may be
shared, i.e. it may cache responses from several servers served to
several clients. Another example is an institutional forward proxy,
which are close to several users. Finally, the User Agent may
implement a private cache for its user at the client side.

@@

The \cite{dyldo2} Dynamic Linked Data Observatory (DyLDO) performed,
and continue to do so, monitoring of parts of the Linked Open Data
Cloud to determine dynamicity characteristics of Linked Data. Caching
is one of their motivations, but they have not published statistics on
HTTP headers.

Linked Data Fragments \cite{ldf1} is said to take advantage of caching
and contrasts this with the unavailability of SPARQL query caches, and
it is claimed that this is an architectural problem.

@@

A general caching approach is given in \cite{yang2011caching}.  In
\cite{umbrich2012hybrid}, the term caching was used in a different
sense than we do, they rather prefetch an entire dataset to a local
store and based on heuristics tried to determine which parts of the
query should be evaluated remotely and locally. \cite{lampo2011cache}
explored when caching had a positive effect on complex SPARQL
queries.

@@

First, we investigated whether certain servers provided better support
for caching than others. To do this, we formulated SPARQL queries to
examine the \httph{Server} headers of successful responses, with
optional clauses matching the standard-compliant computed freshness
time (which is the ideal) as well as whether the response had other
promising features that could assist caching, modification time,
certain predicates, etc.

@@

For a more systematic approach, we wish to test the hypothesis that
some servers are better configured to support caching than others. To
test this hypothesis, we create two so-called \emph{contigency
  tables} (see e.g. \cite{kn:bj}). This formalism is suited to see if the
distribution of  \httph{Server} headers are different for those implementations that
offer caching headers from those that don't. Intuitively, we expect
these distributions to be similar, ``long-tail'' distributions, i.e. a
handful of servers are used by a large number of projects, and then it
falls off rapidly, and so, some servers are used only by very
few. Likewise, it is only to be expected that only a few projects have
given caching enough attention, but that they account for the majority
of the support. The question is if the presence of caching headers is
merely a matter of proportion, or if there are some that have given it
more attention, but still is in relatively little use.

@@

We found that support for HTTP caching and conditional requests are
occurring regularly in the Semantic Web, but they are not
commonplace. We found, in agreement with DyLDO\cite{dyldo2}, that many
resources change at a very slow pace, but also that this is not
reflected in the standards-compliant freshness lifetimes advertized by
servers.

@@

The Billion Triple Challenge 2014 dataset provided data in the form of
NQuad files. Due to the presence of invalid RDF, we iterated through
the NQuad files on a line-by-line basis. First, we matched each line
against a regular expression were lines matching
\texttt{ontology|endpoint|sparql|vocabular} passed the filter. Then,
the Perl framework
RDF::Trine\footnote{https://metacpan.org/release/GWILLIAMS/RDF-Trine-1.011} was used
to parse the line. Lines that failed to parse was discarded. We have
not investigated whether this could introduce biases. Statements were
then accepted into a new NQuad file if they had a predicate that
matched the
\rdfterm{sd:endpoint}\footnote{http://www.w3.org/TR/sparql11-service-description/\#sd-endpoint}
or matched a case-insensitive regular expression \texttt{sparql} if
the subject and objects were both resources, or the predicates
\rdfterm{void:vocabulary}\footnote{http://www.w3.org/TR/void/\#vocabularies},
\rdfterm{rdfa:vocabulary}\footnote{http://www.w3.org/TR/rdfa-core/\#term-prefix-definitions-vocabulary}
or
\rdfterm{api:vocabulary}\footnote{http://purl.org/linked-data/api/vocab\#vocabulary},
as well as having an resource as object. Finally, statements with the
the classes
\rdfterm{cogs:Endpoint}\footnote{http://vocab.deri.ie/cogs\#Endpoint},
\rdfterm{owl:Ontology} and
\rdfterm{voaf:Vocabulary}\footnote{http://lov.okfn.org/vocab/voaf/v2.3/index.html\#Vocabulary}
in the object position were also accepted. More classes and properties
were considered, but not used in the data reduction if they did not
occur in the original data.

>>>>

Caching has been given a prominent place in the foundational documents
of the World Wide Web. Out of the 6 documents that make up the HTTP
1.1 standard, RFC7234 \cite{rfc7234} is entirely devoted to the topic,
and RFC7232 \cite{rfc7232} defines conditional requests, which is also
important when constructing caches. As the former notes:
\begin{quote} 
  The goal of caching in HTTP/1.1 is to significantly improve
  performance by reusing a prior response message to satisfy a current
  request.
\end{quote}
Furthermore, the Architecture of the World Wide Web
\cite{Jacobs:04:AWW} discusses caching throughout, and the definition
of the Representational State Transfer (REST) architectural style
\cite{Fielding_2000_Architectural-Styles} is partly motivated from the
requirement to implement efficient caching. We also note that caching
in the Internet infrastructure, through so-called Content Delivery
Networks, is both a large business area and something that could
provide great value to the Semantic Web.

@@

We note that caching is not only useful for long-living information
resources, even though that may be the most important use. If a resource
is frequently requested, it may make sense to cache it even though it
may be fresh for only a very short period.

Caching may be deployed at several different levels: An HTTP cache may
be in a reverse proxy close to the server, in which case it may have
much in common with a conventional database cache. It may also be
anywhere between a server and a client, in which case it may be
shared, i.e. it may cache responses from several servers served to
several clients. Another example is an institutional forward proxy,
which are close to several users. Finally, the User Agent may
implement a private cache for its user at the client side.

@@

The \cite{dyldo2} Dynamic Linked Data Observatory (DyLDO) performed,
and continues to do so, monitoring of parts of the Linked Open Data
Cloud to determine dynamicity characteristics of Linked Data. Caching
is one of their motivations, but they have not published statistics on
HTTP headers.

Linked Data Fragments is claimed in \cite{ldf1} to take advantage of
caching and contrasts this with the unavailability of SPARQL query
caches, and it is claimed that this is an architectural problem.

@@

A general caching approach is given in \cite{yang2011caching}.  In
\cite{umbrich2012hybrid}, the term caching was used in a different
sense than we do, they rather prefetched an entire dataset to a local
store and based on heuristics tried to determine which parts of the
query should be evaluated remotely and locally. \cite{lampo2011cache}
explored when caching had a positive effect on complex SPARQL
queries.

@@

First, we investigated whether certain server implementations provided better support
for caching than others. To do this, we formulated SPARQL queries to
examine the \httph{Server} headers of successful responses, with
optional clauses matching the standard-compliant computed freshness
time (which is the ideal) as well as whether the response had other
promising features that could assist caching, modification time,
certain predicates, etc.

@@

For a more systematic approach, we wish to test the hypothesis that
some servers are better configured to support caching than others. To
test this hypothesis, we create two so-called \emph{contigency
  tables} (see e.g. \cite{kn:bj}). This formalism is suited to see if the
distribution of  \httph{Server} headers are different for those implementations that
offer caching headers from those that don't. Intuitively, we expect
these distributions to be similar, ``long-tail'' distributions, i.e. a
handful of servers are used by a large number of projects, and then it
falls off rapidly, and so, some servers are used only by very
few. Likewise, it is to be expected that only a few projects have
given caching enough attention, but that they account for the majority
of the support. The question is if the presence of caching headers is
merely a matter of proportion, or if there are some that have given it
more attention, but still is in relatively little use.

@@

We found moderate uptake for HTTP caching and conditional requests in
the Semantic Web. We found, in agreement with DyLDO\cite{dyldo2}, that
many resources change at a very slow pace, but also that this is not
reflected in the standards-compliant freshness lifetimes advertized by
servers.

@@

The Billion Triple Challenge 2014 dataset provided data in the form of
NQuad files. Due to the presence of invalid RDF, we iterated through
the NQuad files on a line-by-line basis. First, we matched each line
against a regular expression were lines matching
\texttt{ontology|endpoint|sparql|vocabular} passed the filter. Then,
the Perl framework
RDF::Trine\footnote{https://metacpan.org/release/GWILLIAMS/RDF-Trine-1.011} was used
to parse the line. Lines that failed to parse was discarded. We have
not investigated whether this could introduce biases. Statements were
then accepted into a new NQuad file if they had a predicate that
matched the
\rdfterm{sd:endpoint}\footnote{http://www.w3.org/TR/sparql11-service-description/\#sd-endpoint}
or matched a case-insensitive regular expression \texttt{sparql} if
the subject and objects were both resources, or the predicates
\rdfterm{void:vocabulary}\footnote{http://www.w3.org/TR/void/\#vocabularies},
\rdfterm{rdfa:vocabulary}\footnote{http://www.w3.org/TR/rdfa-core/\#term-prefix-definitions-vocabulary}
or
\rdfterm{api:vocabulary}\footnote{http://purl.org/linked-data/api/vocab\#vocabulary},
as well as having a resource as object. Finally, statements with the
the classes
\rdfterm{cogs:Endpoint}\footnote{http://vocab.deri.ie/cogs\#Endpoint},
\rdfterm{owl:Ontology} and
\rdfterm{voaf:Vocabulary}\footnote{http://lov.okfn.org/vocab/voaf/v2.3/index.html\#Vocabulary}
in the object position were also accepted. More classes and properties
were considered, but not used in the data reduction if they did not
occur in the original data.
